Simulations of Ballot Polling Risk-Limiting Audits

### 2021-10-31 MultipleRoundSims/draft.tex
added material on R2B2
changed definition of stopping prob
bib file, changes to intro, reordered related work and defs
Related work
Done with a first edit

% This is a draft of the coming paper "Multiple Round Ballot Polling Risk-Limiting Audit Simulations".
% 2021.
%
\documentclass[runningheads]{llncs}
%
\usepackage{float}
\usepackage{xspace}
\usepackage{graphicx}
\graphicspath{{./imgs/}}
\usepackage{comment}

@@

\begin{document}
%
\title{Simulations of Ballot Polling Risk-Limiting Audits}
%add \thanks{} within the title brackets if want to refer to supporting organization ^^
%
%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\begin{comment}
\author{Oliver Broadrick\inst{1} \and
Sarah Morin\inst{1} \and
Grant McClearn\inst{2} \and Neal McBurnett \and Poorvi L. Vora\inst{1} \and
Filip Zagorski\inst{3} }
%
\authorrunning{Broadrick, Morin et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Department of Computer Science, The George Washington University 
\email{obroadrick@gwmail.gwu.edu}\\
\and
Department of Computer Science, Stanford University 
\email{grantmcc@stanford.edu}\\
\and Wroclaw University of Science and Technology\\
\email{filip.zagorski@gmail.com}}
%
\end{comment}
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
    In this paper we present simulation results comparing the risk, stopping probability and number of ballots required over multiple rounds of ballot-polling risk limiting audits (RLAs) \Minerva, Selection-Ordered \BRAVO and End-of-Round \BRAVO. We also
    present details on the R2B2 open source software library and the related simulation software.

@@

    An open question, however, is how \Minerva compares to Selection-Ordered \BRAVO and End-of-Round \BRAVO over multiple rounds.  In this paper, we present results from simulations of multiple round audits with first-round stopping probabilities of $0.9$, a common choice among election officials. Because the size of rounds in \Minerva needs to be predetermined (and independent of the audit draws), we use two pre-determined round sequences for \Minerva: (a) all rounds are of the same size (which might be a practical choice for election officials if they had planned to have resources for the first round size) and (b) each round is $1.5$ times the previous one (which is the preset value for \Minerva as integrated into election audit software Arlo).

@@

\section{Definitions}
\label{sec:defs}
\input{sections/02-defs}

\section{Related work}
\label{sec:related}
\input{sections/03-related}

\section{Simulator}
\label{sec:simulator}
\input{sections/04-simulator}

@@

%
% ---- Bibliography ----
%
% BibTeX users should specify bibliography style 'splncs04'.
% References will then be sorted and formatted in the correct style.
%
% \bibliographystyle{splncs04}
% \bibliography{mybibliography}
%
\begin{thebibliography}{8}
\bibitem{ref_article1}
Author, F.: Article title. Journal \textbf{2}(5), 99--110 (2016)

@@

\bibitem{ref_url1}
LNCS Homepage, \url{http://www.springer.com/lncs}. Last accessed 4
Oct 2017
\end{thebibliography}
\end{document}

>>>>

% This is a draft of the coming paper "Multiple Round Ballot Polling Risk-Limiting Audit Simulations".
% 2021.
%
\documentclass[runningheads]{llncs}
%
\usepackage{float}
\usepackage{xspace}
\usepackage{graphicx}
\graphicspath{{./imgs/}}
\usepackage{comment}
\usepackage{listings,hyperref}

@@

\def\code#1{\texttt{#1}}

\begin{document}
\lstset{language=Python}
%
\title{Simulations of Ballot Polling Risk-Limiting Audits}
%add \thanks{} within the title brackets if want to refer to supporting organization ^^
%
%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\begin{comment}
\author{Oliver Broadrick\inst{1} \and
Sarah Morin\inst{1} \and
Grant McClearn\inst{2} \and Neal McBurnett \and Poorvi L. Vora\inst{1} \and
Filip Zag{\'o}rski\inst{3} }
%
\authorrunning{Broadrick, Morin et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Department of Computer Science, The George Washington University 
\email{obroadrick@gwmail.gwu.edu}\\
\and
Department of Computer Science, Stanford University 
\email{grantmcc@stanford.edu}\\
\and Wroclaw University of Science and Technology\\
\email{filip.zagorski@gmail.com}}
%
\end{comment}
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
    In this paper we present simulation results comparing the risk, stopping probability and number of ballots required over multiple rounds of ballot-polling risk limiting audits (RLAs) \Minerva, Selection-Ordered \BRAVO and End-of-Round \BRAVO. We also
    present details on the R2B2 open source software library and the related simulation software.

@@

    An open question, however, is how \Minerva compares to Selection-Ordered \BRAVO and End-of-Round \BRAVO over multiple rounds.  In this paper, we present results from simulations of multiple round audits with first-round stopping probabilities of $90\%$, a common choice among election officials. Because the size of rounds in \Minerva needs to be predetermined (and independent of the audit draws), we use two pre-determined round sequences for \Minerva: (a) all rounds are of the same size (which might be a practical choice for election officials if they had planned to have resources for the first round size) and (b) each round is $1.5$ times the previous one (which is the preset value for \Minerva as integrated into election audit software Arlo).

@@

\section{Related work}
\label{sec:related}
\input{sections/02-related}

\section{Software}
\label{sec:software}
\input{sections/03-software}

\section{Experiments}
\label{sec:expts}
\input{sections/04-expts}

@@

%
% ---- Bibliography ----
%
% BibTeX users should specify bibliography style 'splncs04'.
% References will then be sorted and formatted in the correct style.
%
\bibliographystyle{splncs04}
\bibliography{audits}
%
\end{document}

\begin{thebibliography}{8}
\bibitem{ref_article1}
Author, F.: Article title. Journal \textbf{2}(5), 99--110 (2016)

@@

\bibitem{ref_url1}
LNCS Homepage, \url{http://www.springer.com/lncs}. Last accessed 4
Oct 2017
\end{thebibliography}

### 2021-10-31 MultipleRoundSims/sections/03-software.tex
Done with a first edit

The R2B2 Python library \cite{r2b2} provides a framework for the exploration of round-by-round
and ballot-by-ballot RLAs. The goal in designing R2B2 is two fold:
\begin{enumerate}
    \item Provide an elegant Python library which can be easily imported and used
    in any other code base.
    \item Provide an interactive set of tools which can be utilized `out-of-the-box'
    for experimenting with and learning about risk-limiting audits.
\end{enumerate}

@@

\subsection{Simulation Software}
As described above, our open source audit software library R2B2 has implementations of several ballot polling risk-limiting audits as well as a simulator, 
all written in Python.
For each of these audits, the software can compute the stopping condition for a given sample and estimates
of the next round size to achieve a desired stopping probability. 
For a given audit and random seed, the simulator draws random samples, with replacement, using a pseudorandom number generator, [need to check].
given the number of votes for each candidate, and the number of invalid votes, in the underlying election (these need not be chosen to be those announced).

When the number of candidates is more than two, the audit is carried out pairwise for each candidate pair, and votes for all other candidates are considered invalid votes. 
It is often useful to consider two underlying elections: one corresponding to a tie between the pair, and the other corresponding to the results as announced. These are the probability distributions represented
by the null and alternative hypotheses.

@@

The R2B2 (python) software is used to compute stopping conditions and next round estimates, it is intended for use by us and other researchers, and designed for this purpose. We have also independently implemented all the functionality in matlab (the two codebases are written by different individuals) and have extensively checked the results of the two codebases. Additionally, for use in regular election audits by election officials, we have written an add-on to the {\em Arlo} risk-limiting audit software, which has also been extensively checked against these two codebases.

>>>>

The R2B2 Python library \cite{r2b2_anon} provides a framework for the exploration of round-by-round
and ballot-by-ballot RLAs. The goal in designing R2B2 is two fold:
\begin{enumerate}
    \item Provide an elegant Python library which can be easily imported and used
    in any other code base.
    \item Provide an interactive set of tools which can be utilized `out-of-the-box'
    for experimenting with and learning about risk-limiting audits.
\end{enumerate}

@@

\subsection{Simulation Software}
As described above R2B2 has implementations of several ballot polling risk-limiting audits as well as a simulator, 
all written in Python.
For each of these audits, the software can compute the stopping condition for a given sample and estimates
of the next round size to achieve a desired stopping probability. 
For a given audit and random seed, the simulator draws random samples, with replacement, using a pseudorandom number generator, [need to check].
given the number of votes for each candidate, and the number of invalid votes, in the underlying election (these need not be chosen to be those announced).

When the number of candidates is more than two, the audit is carried out pairwise for each candidate pair, and votes for all other candidates are considered invalid votes.

@@

The R2B2 software is used to compute stopping conditions and next round estimates. It is intended for use by us and other researchers, and designed for this purpose. We have also independently implemented all the functionality in matlab \cite{brla_explore_anon} (the two codebases are written by different individuals) and have extensively checked the results of the two codebases. Additionally, for use in regular election audits by election officials, we have written an add-on \cite{athena_anon} to the {\em Arlo} risk-limiting audit software, the results of which have also been extensively checked against the other two codebases.

### 2021-10-31 MultipleRoundSims/sections/04-expts.tex
Done with a first edit

%Definitions
Now we present relevant definitions.
We consider a two candidate plurality contest.
To begin we define an audit. We assume that ballots are sampled with replacement. 
\begin{definition}
An audit $\mathcal{A}$ takes a sample $X$ as input and gives one of the 
following decisions
\begin{enumerate}
\item
$Correct$: the audit is complete
\item
$Uncertain$: continue the audit
\end{enumerate}
\end{definition}
All of the audits discussed in this paper are modeled as binary hypothesis tests.
Under the alternate hypothesis, $H_a$, the announced outcome is correct. 
That is, the true underlying ballot distribution is given by the announced ballot tallies.
Under the null hypothesis, $H_0$, the true outcome is a tie 
(or the announced winner lost by one vote, and the number of ballots is large enough that the probability of drawing a ballot for the winner is that of drawing one for the winner).
Now we define key attributes of an audit that we will consider while analyzing our simulations.
The maximum risk of an audit is the probability that an audit stops, given that the underlying election is a tie. 
\begin{definition}[Risk]
The maximum risk $R$ of an audit $\mathcal{A}$ is
$$R(\mathcal{A})=\Pr[\mathcal{A}(X)=Correct \mid H_0]$$
\end{definition}
This leads us to the following simple definition of an $\alpha$-RLA.
\begin{definition}[Risk Limiting Audit ($\alpha$-RLA)]
An audit $\mathcal{A}$ is a Risk Limiting Audit with 
risk limit $\alpha$ iff 
$$R(\mathcal{A}) \le \alpha.$$
\end{definition}

@@

>>>>

%Definitions
In this section, we motivate and describe the experiments. We consider a two candidate plurality contest, and assume that ballots are sampled with replacement, as is common in the literature.

We first present relevant definitions.

\begin{definition}
An audit $\mathcal{A}$ takes a sample of ballots $X$ as input and gives one of the 
following decisions
\begin{enumerate}
\item
$Correct$: the audit is complete
\item
$Uncertain$: continue the audit
\end{enumerate}
\end{definition}

All of the audits discussed in this paper are modeled as binary hypothesis tests.

Under the alternate hypothesis, $H_a$, the announced outcome is correct. 
That is, the true underlying ballot distribution is given by the announced ballot tallies.

Under the null hypothesis, $H_0$, the true outcome is a tie 
(or the announced winner lost by one vote, and the number of ballots is large enough that the probability of drawing a ballot for the winner is that of drawing one for the winner).

The maximum risk of an audit is the probability that an audit stops, given that the underlying election is a tie. (Vora show that this is the maximum risk \cite{Bayesian-RLA}.) 
\begin{definition}[Risk]
The maximum risk $R$ of an audit $\mathcal{A}$ is
$$R(\mathcal{A})=\Pr[\mathcal{A}(X)=Correct \mid H_0]$$
\end{definition}

This leads us to the following definition of an $\alpha$-RLA.
\begin{definition}[Risk Limiting Audit ($\alpha$-RLA)]
An audit $\mathcal{A}$ is a Risk Limiting Audit with 
risk limit $\alpha$ iff 
$$R(\mathcal{A}) \le \alpha.$$
\end{definition}

@@

We performed simulations to study $S_j$ and obtain estimates of the expected number of ballots drawn for the various audits. We know that \BRAVO uses the smallest expected number of ballots when drawn ballot-by-ballot. However, this may not be the case when \BRAVO is used as an R2 audit for a large first round size. We also studied the risk values. While all the audits we studied have been proven to be RLAs, it is useful to observe how close to the risk limit each audit gets over a small, finite number of rounds as would be the case in a real election.

### 2021-10-31 MultipleRoundSims/sections/05-results.tex
Done with a first edit

%Results
For this paper, we simulated audits for 
the 2020 Presidential election
in all US states whose margin was at least $5\%$.
Round sizes increase roughly proportional to the inverse
square of the margin, so 
smaller margins are computationally much more expensive to simulate.
For each of these states, we simulated 
$10,000=10^4$ audits with the announced
underyling ballot distribution
and an additional $10,000=10^4$ audits with a tie
as the underlying ballot distribution.
These are reasonable choices for inital simulation experiments
because most audits frame the stopping decision as a binary
hypothesis test where the null hypothesis assumes an underyling tie
and the alternative hypothesis assumes the announced distribution.
A standard first round size in ballot-polling audits
has been one which achieves a $90\%$ probability
of stopping, and we chose round sizes to reflect this standard
in our simulations.
We ran our simulations for up to five rounds.
With a $90\%$ stopping probability, 
an audit only has a $a$ probablity of not stopping
by the end of the fifth round, assuming the outcome was correctly
announced.

\subsection{End-of-Round \BRAVO}
For the EoR \BRAVO simulations, our software estimated and used for each round
the minimum round size that would achieve a $90\%$ probability of stopping.
For the simulations with the announced outcome as the underlying
ballot distribution, for each round $j$ we computed the proportion of audits 
that stopped in round $j$ among those that had not stopped before round $j$.

\begin{figure}[H]
\includegraphics[width=\textwidth]{eor_bravo_90perc_10^4_corrected/sprob_first_three.png}\caption{
For the first three rounds of the EoR \BRAVO simulations, this plot shows the proportion of audits that stopped in the $j$th round
to all audits which had not yet stopped before the $j$th round.}
\label{fig:eor_bravo_sprob}
\end{figure}

In Figure \ref{fig:eor_bravo_sprob}, we display proportions for only the first three rounds
since very few audits, $(.1)^{j-1}\cdot(10^4)$ on average, 
make it to the $jth$ round.
As  a result, the proportions are based on an exponentially smaller dataset in each round.
The proportions shown in Figure \ref{fig:eor_bravo_sprob} give an estimate
of the true probability of an EOR \BRAVO audit stopping in the $j$th round,
given that it has not already stopped in a previous round. 
In \ref{fig:eor_bravo_sprob}, we see that, especially in earlier rounds for which 
the values are more representative of true audit behavior, 
our predictions are accurate.
In particular, the average across all margins is just above $.9=90\%$ for
all three rounds.

The risk of this audit, across all $5$ rounds, is an important metric since it determines whether an audit is risk-limiting.
Therefore, we now consider the proportion of audits that stopped with an underlying tie.
This proportion, for a risk-limiting audit, should approach a value less than the risk limit, $0.1$, as more audits are performed.

\begin{figure}[H]
\includegraphics[width=\textwidth]{eor_bravo_90perc_10^4_corrected/total_risk.png}
\caption{For each state margin, this plot shows
the proportion of EOR \BRAVO audits with an underlying
tie that stopped.}
\label{fig:eor_bravo_risk}
\end{figure}

Figure~\ref{fig:eor_bravo_risk} makes it clear that the risk of EOR \BRAVO is roughly
an order of magnitude less than the risk limit. 
That is, on average, roughly $10$ times more audits could have stopped 
with the audit still meeting the risk limit.
These simulations support the claim made in [\Minerva paper]
that EOR \BRAVO is unnecessarily conservative and thus requires
more ballots on average.

% \subsection{Selection-Ordered BRAVO}
% next_sample_size code is slow... tie simulations running still!

\subsection{\Minerva Simulations}
For \Minerva, it has not been shown that round sizes can be chosen
during the audit. That is, an adversary with knowledge of the history
of the audit may be able to choose round sizes which cause the 
risk of the audit to exceed the risk limit.
For this reason, we have to choose the round sizes of a \Minerva 
audit a priori.
For this paper, we consider two choices of round sizes.
For both, we estimate and then use the minimum first round size 
which achieves
a $90\%$ probability of stopping.
Then, for subsequent rounds, we either (i) 
draw the same number of ballots in each round or (ii)
multiply the previous round size by a factor of $1.5$ and 
sample this many new ballots.
We consider the case of drawing samples of the same size
because it may reflect a practical way to continue an
audit; if election officials have selected some first round size within
reasonable logistical bounds, drawing the same number of 
ballots in subsequent rounds may be practical.
We also consider round sizes with samples increasing by a multiple
of $1.5$ because this multiple gives a very rough approximation of 
round sizes with a $90\%$ probability of stopping.

\subsubsection{Round Sizes with Multiple of $1.0$}

As with the preceding simulations, we ran $10,000=10^4$ trials
per state for both the underlying tie and underlying reported
outcome.

\begin{figure}[H]
\includegraphics[width=\textwidth]{minerva_multiround_1x_10^4/total_risk.png}
\caption{This plot shows, for each state's margin, the proportion of audits that stopped of
all $10^4$ audits with an underlying tie in the simulations with a round size multiple of $1.0$.}
\label{fig:minerva1_risk}
\end{figure}

Notice in Figure \ref{fig:minerva1_risk} that all simulations for a tie had a proportion of audits that stopped less than $.1$, the risk limit, supporting
the claim that \Minerva is a risk limiting audit. 
Unlike EOR \BRAVO, the experimental risks here are much closer to the risk limit,
showing that \Minerva stops on average with a less conservative risk; \Minerva is sharper.

\begin{figure}[H]
\includegraphics[width=\textwidth]{minerva_multiround_1x_10^4/sprobs_first_three.png}
\caption{This plot shows, for each state's margin, the proportion of audits that stopped of
all $10^4$ audits with the announced results as the underlying distribution
in the \Minerva simulations with a round size multiple of $1.0$.}
\label{fig:minerva1_sprob}
\end{figure}

Figure~\ref{fig:minerva1_sprob} shows that the round size estimate for the first round achieved the desired
stopping probability of $90\%$. For subsequent rounds, the multiplier of $1.5$ did not consistently achieve $90\%$ stopping
probability, but it was within  roughly $10\%$.

\subsubsection{Round Sizes with Multiple of $1.5$}

@@

\begin{figure}[H]
\includegraphics[width=\textwidth]{minerva_multiround_1p5x_10^4/total_risk.png}
\caption{This plot shows, for each state's margin, the proportion of audits that stopped of
all $10^4$ audits with an underlying tie in the simulations with a round size multiple of $1.5$.}
\label{fig:minerva1p5_risk}
\end{figure}

Figure~\ref{fig:minerva1p5_risk} shows, for each state's margin, the proportion of audits that stopped of
all $10^4$ audits with an underlying tie in the simulations with a round size multiple of $1.5$.

\begin{figure}[H]
\includegraphics[width=\textwidth]{minerva_multiround_1p5x_10^4/sprobs_first_three.png}
\caption{This plot shows, for each state's margin, the proportion of audits that stopped in each round
with the announced results as the underlying distribution
in the \Minerva simulations with a round size multiple of $1.5$.}
\label{fig:minerva1p5_sprob}
\end{figure}

Figure~\ref{fig:minerva1p5_sprob} shows that the round size estimate for the first round achieved the desired
stopping probability of $90\%$. For subsequent rounds, the multiplier of $1.5$ did not consistently achieve $90\%$ stopping
probability, but it was within  roughly $10\%$. 
% we could suggest that predicting accurate multiples is possible (same curve rather than line)

>>>>

%Results
For this paper, we simulated audits for 
the 2020 Presidential election
in all US states where the pairwise margin for the two main candidates was at least $5\%$. 
Round sizes increase roughly proportional to the inverse
square of the margin, so 
smaller margins are computationally much more expensive to simulate.
For each of these states, we simulated 
$10,000=10^4$ audits assuming the underlying election was as announced,  
and an additional $10,000=10^4$ audits assuming the underlying election was a tie.

We ran simulations for a $90\%$ probability of stopping in the first round, enabling election officials to be done in the first round with very high probability if the election is as announced.
We ran our simulations for up to five rounds. We assume that the audit results in a hand count if it does not stop in any of $5$ rounds.

\subsection{End-of-Round \BRAVO}
For the EoR \BRAVO simulations, our software estimated and used for each round
the round size that would achieve a $90\%$ probability of stopping if the true election outcome were as announced. In Figure \ref{fig:eor_bravo_sprob}, we display the proportion of EoR \BRAVO audits that stopped in the $j^{th}$ round
to all audits which had not yet stopped before the $j^{th}$ round, for only the first three rounds of the simulations because very few audits, $(.1)^{j-1}\cdot(10^4)$ on average, 
make it to the $j^{th}$ round. The proportions give an estimate
of the true probability of an EoR \BRAVO audit stopping in the $j$th round,
given that it has not already stopped in a previous round, when the underlying election is as announced. 
We see that, especially in earlier rounds for which 
the values are more representative of true audit behavior, 
our round size predictions are accurate.
In particular, the average across all margins is just above $.9=90\%$ for
all three rounds.

We also study the proportion of audits that stopped when the underlying election was a tie.
This proportion should approach a value less than the risk limit, $0.1$, as more audits are performed.

\begin{figure}
\includegraphics[width=0.8\textwidth]{eor_bravo_90perc_10^4_corrected/sprob_first_three_cropped.png}\caption{
This plot shows, for each state margin, when the underlying election is as announced, the number of EoR \BRAVO audits that stopped in the $j^{th}$ round,
as a fraction of all EoR \BRAVO audits which had not yet stopped before the $j^{th}$ round for $j=1,2,3$.}
\label{fig:eor_bravo_sprob}
\end{figure}

\begin{figure}
\includegraphics[width=0.8\textwidth]{eor_bravo_90perc_10^4_corrected/total_risk.png}
\caption{This plot shows, for each state margin,
the fraction of EoR \BRAVO audits that stopped in any of the $5$ rounds when the underlying election was a tie.}
\label{fig:eor_bravo_risk}
\end{figure}

We observe (see Figure~\ref{fig:eor_bravo_risk}) that the risk of EoR \BRAVO is roughly
an order of magnitude less than the risk limit. 
These results are as expected, because EoR \BRAVO is known to be too conservative \cite{usenix_minerva}.

\subsection{Selection-Ordered \BRAVO}
% next_sample_size code is slow... tie simulations running still!
For the SO \BRAVO simulations, our software estimated and used for each round
the round size that would achieve a $90\%$ probability of stopping.

In Figure \ref{fig:so_bravo_sprob}, we again display proportions of SO \BRAVO audits that stopped in the $j^{th}$ round
to all audits which had not yet stopped before the $j^{th}$ round for only the first three rounds.
The proportions shown in Figure \ref{fig:so_bravo_sprob}, like those in \ref{fig:eor_bravo_sprob}, give an estimate
of the true probability of an SO \BRAVO audit stopping in the $j^{th}$ round,
given that it has not already stopped in a previous round. 
In \ref{fig:so_bravo_sprob}, we see that our round size predictions are relatively accurate,
all three rounds being near $90\%$.

\begin{figure}
\includegraphics[width=0.8\textwidth]{so_bravo_90perc_10^4/sprob_first_three.png}\caption{
This plot shows, for each state margin, when the underlying election is as announced, the number of SO \BRAVO audits that stopped in the $j^{th}$ round,
as a fraction of all EoR \BRAVO audits which had not yet stopped before the $j^{th}$ round for $j=1,2,3$.}
\label{fig:so_bravo_sprob}
\end{figure}

As before, we will next consider the proportion of audits that stopped with an underlying tie over all five rounds.
This proportion, for a risk-limiting audit, should approach a value less than the risk limit, 
$0.1$, as more audits are performed.

In Figure~\ref{fig:so_bravo_risk} we show only the results for the $13$
states whose simulations with an underlying tie have finished running.
To estimate the next round size that achieves a desired stopping probability,
the SO \BRAVO software generates the probability distribution on the number of ballots in the sample ballot by ballot (see \cite{usenix_minerva}) since
the stopping condition needs to be evaluated for each individual ballot drawn.
because the underlying tied election causes audits to move on to larger rounds, 
the simulations are computationally expensive. SO \BRAVO is proven to be a Risk-Limiting Audit,
and we observe in Figure~\ref{fig:so_bravo_risk},
that the risk of SO \BRAVO is much
nearer the risk limit than that of EoR \BRAVO, as expected.

\begin{figure}
\includegraphics[width=0.8\textwidth]{so_bravo_90perc_10^4/total_risk.png}
\caption{This plot shows, for each state margin,
the fraction of SO \BRAVO audits that stopped in any of the $5$ rounds when the underlying election was a tie.}
\label{fig:so_bravo_risk}
\end{figure}

%TODO SO and EoR macros

\subsection{\Minerva Simulations}
The proof that \Minerva is an RLA \cite{usenix_minerva} assumes that the round schedule is pre-determined. For this reason, we have to choose the round sizes of a \Minerva 
audit {\em a priori}.
For this paper, we consider two choices of round sizes.
For both, we estimate and use a first round size 
for
a $90\%$ probability of stopping.
Then, for subsequent rounds, we either (i) 
draw the same number of ballots in each round or (ii)
multiply the previous round size by a factor of $1.5$ and 
sample this many new ballots.
We consider the case of drawing samples of the same size
because it may reflect a practical way to continue an
audit; if election officials have selected some first round size within
reasonable logistical bounds, drawing the same number of 
ballots in subsequent rounds may be practical.
We also consider round sizes with samples increasing by a multiple
of $1.5$ because this multiple gives a very rough approximation of 
round sizes with a $90\%$ probability of stopping for \Minerva.

As with the preceding simulations, we ran $10,000=10^4$ trials
per state for both an underlying election as reported and an underlying tied election.

Figure~\ref{fig:minerva1_sprob} and Figure~\ref{fig:minerva1p5_sprob} show that the first round size estimates were fairly accurate for multipliers of $10.$ and $1.5$ respectively, with first round stopping probabilities being very close to $90\%$. For subsequent rounds, the multipliers of $1.0$ and $1.5$ respectively did not consistently achieve $90\%$ stopping
probability, as they were not accurate estimates of $90\%$ stopping probabilities. Note that we chose a simple multiplier for future rounds, but one could make more accurate round size estimates before the audit begins.

\begin{figure}
\includegraphics[width=0.8\textwidth]{minerva_multiround_1x_10^4/sprobs_first_three.png}
\caption{This plot shows, for each state margin, when the underlying election is as announced, the number of \Minerva audits that stopped in the $j^{th}$ round,
as a fraction of all \Minerva audits which had not yet stopped before the $j^{th}$ round for $j=1,2,3$ and round size multiple of $1.0$.}
\label{fig:minerva1_sprob}
\end{figure}

\begin{figure}
\includegraphics[width=0.8\textwidth]{minerva_multiround_1p5x_10^4/sprobs_first_three.png}
\caption{This plot shows, for each state margin, when the underlying election is as announced, the number of \Minerva audits that stopped in the $j^{th}$ round,
as a fraction of all \Minerva audits which had not yet stopped before the $j^{th}$ round for $j=1,2,3$ and round size multiple of $1.5$.}
\label{fig:minerva1p5_sprob}
\end{figure}

Figures \ref{fig:minerva1_risk} and Figure~\ref{fig:minerva1p5_risk} show that fewer than $0.1$ of the audits stopped when the underlying election was a tie, for round multiples of $1.0$ and $1.5$ respectively, as would be expected for an RLA with risk limit $0.1$. 
Unlike EOR \BRAVO, the experimental risks here are much closer to the risk limit,
showing that \Minerva stops on average with a less conservative risk; \Minerva is sharper.

\begin{figure}
\includegraphics[width=0.8\textwidth]{minerva_multiround_1x_10^4/total_risk.png}
\caption{This plot shows, for each state margin,
the fraction of \Minerva audits with a round size multiple of $1.0$ that stopped in any of the $5$ rounds when the underlying election was a tie.}
\label{fig:minerva1_risk}
\end{figure}

@@

% we could suggest that predicting accurate multiples is possible (same curve rather than line)

\begin{figure}
\includegraphics[width=0.8\textwidth]{minerva_multiround_1p5x_10^4/total_risk.png}
\caption{This plot shows, for each state margin,
the fraction of \Minerva audits with a round size multiple of $1.5$ that stopped in any of the $5$ rounds when the underlying election was a tie.}
\label{fig:minerva1p5_risk}
\end{figure}
